### 2017-07-03 reasoning.tex
Edits.

Two rules of the transaction-local RG judgment are shown in
Fig.~\ref{fig:rg-rules}. The rule \rulelabel{RG-Update} is
illustrative of the RG rules for SQL statements; they basically
reflect the structure of the corresponding reduction rule from
Fig.~\ref{fig:txnimp}, and require no further explanation. The rule
\rulelabel{RG-Foreach} defines the RG judgment for a \C{FOREACH} loop.
As is characteristic of loops, the reasoning is pivoted on a loop
invariant $\I$ (not same as the high-level invariant), that needs to
be stable w.r.t $\R$. $I$ must follow $P$, the pre-condition of
\C{FOREACH}, when no elements have been iterated, i.e, when
$y=\emptyset$. The body of the loop can assume the loop invariant, and
the fact that $z$ is an element from the set $x$ being iterated, to
prove its post-condition $Q_c$. Operational semantics ensures that $z$
is added to $y$ at the end of the iteration, hence $Q_c \conj z\in y$
is valid at the end of the iteration, from which $I$ must follow. When
\C{FOREACH} has finished execution, $y$, the set of iterated items, is
the entire set $x$. Thus $I \conj y=x$ must imply the post-condition
$Q$, which also needs to be stable. The rules for conditions,
sequencing etc., are more-or-less standard, hence elided.

% \subsection{Semantics and Soundness}

% The rules that define rely-guarantee judgment are shown in
% Fig.~\ref{fig:rg-rules}. Like standard rely-guarantee definitions,
% these definitions also require a \emph{stability} condition, which
% requires pre- and post- conditions to hold despite any interference
% from concurrent threads (captured by $R$). Stability can be predicated
% on the assumption that interference preserves the trace invariant
% $\I$. Formally:\vspace*{-10pt}

% \begin{smathpar}
% \begin{array}{lcl}
%   \underI{\stable(R,P)} & \defeq & \forall \E, \E'.\, 
%   \I(\E) \conj P(\E) \conj R(\E,\E') \\
%   &   & \hspace*{1in}\conj \I(\E') \Rightarrow P(\E')\\
% \end{array}
% \end{smathpar}

% \noindent However, the assumption that interference preserves $\I$ (or,
% dually, $\I$ withstands interference) needs to be justified
% separately. We call this the stability requirement on $\I$:\vspace*{-10pt}

% \begin{smathpar}
% \begin{array}{lcl}
% \stable(R,\I)& \defeq & \forall \E, \E'.\, 
%   \I(\E) \conj R(\E,\E')\Rightarrow \I(\E')\\
% \end{array}
% \end{smathpar}

% \noindent The rule \rulelabel{RG-Var} defines the rely-guarantee judgment for
% shared variable reads inside a transaction $T_i$. It requires $\I$ to
% be stable, and pre- and post- conditions to be stable relative to
% $\I$.  The quantified premise effectively requires a proof that if the
% abstract machine of Fig.~\ref{fig:txnimp} takes a step starting from
% an execution $\E$ that satisfies the pre-condition $P$, then the
% resultant execution $\E'$ satisfies the post-condition $Q$, and that
% the guarantee $G$ faithfully captures the transition from $\E$ to
% $\E'$. It is informative to compare this premise with the premise of
% the \rulelabel{E-Aux} reduction rule of Fig.~\ref{fig:txnimp}. Similar
% premises also appear in the \rulelabel{RG-Asgn} and \rulelabel{RG-Txn}
% rules, which define rely-guarantee judgments for assignments and
% transactions, respectively. \rulelabel{RG-Var} however also requires
% the return value ($\interp{S}(X)$) of the read to satisfy the
% assertion $C$ meant for the value. \rulelabel{RG-Arith} defines the RG
% judgment for an arithmetic expression $e_1\pm e_2$ in terms of the
% corresponding judgments for the constituent expressions $e_1$ and
% $e_2$. The quantified premise requires any value resulting from
% evaluating $e_1 \pm e_2$ to satisfy the assertion $C$, provided that
% $e_1$ and $e_2$ always evaluate to values that satisfy $C_1$ and
% $C_2$, respectively. The rules for sequential and parallel composition
% of commands are essentially the same as their counterparts in a
% standard rely-guarantee formulation and hence elided.

% % RG judgment for the assignment $X:=e$ makes use of the corresponding
% % judgment for the RHS expression $e$. The quantified premise asserts
% % that evaluating the assigment after evaluating $e$ to a value $v$ and
% % an execution $\E$, should result in an execution $\E'$ that satisfies
% % the post-condition $Q$, while being related to $\E$ via the guarantee
% % $G$. RG judgment for the \C{txn} lexical block is similar.  It uses
% % the judgment for the transaction-bound command $c$ (i.e.,
% % $\txnbox{c}_i$) to obtain an invariant $Q'$ for the execution $\E$
% % before the commit, and verifies that committing the transaction under
% % $\E$ results in an execution $\E'$ that satisfies transaction's
% % post-condition $Q$. As usual, $\E$ and $\E'$ need to be related by $G$.
% % The rules for sequential composition of transaction-bound comands
% % (\rulelabel{RG-Seq}) and parallel composition of top-level commands
% % (\rulelabel{RG-Par}) are straightforward, and more-or-less same as the
% % corresponding rules in classical rely-guarantee.

% The \rulelabel{RG-Conseq} rule defines ways to strengthen or
% weaken relations and assertions associated with the RG judgment of
% transaction-bound expressions. Similar rules exist for
% transaction-bound and top-level commands, but are not
% shown.\footnote{The supplementary materials provide the complete set
%   of rules.} As is the case with a standard rely-guarantee
% formulation,the rules allow the pre-condition $P$ and the rely relation
% $R$ to be strengthened, and the post-condition $Q$ (also, $C$ in the
% case of expressions) and the guarantee relation $G$ to be weakened.
% The most notable aspect of the \rulelabel{RG-Conseq} rules is that
% they allow the trace invariant $\I$ to be strengthened. Considering
% that $\I$ captures isolation properties, this means that a program
% proven correct under weaker isolation levels is also correct under
% stronger ones.  Parametricity over the trace invariant $\I$, combined
% with the ability to strengthen $\I$ as needed, allows our proof system
% to support a highly flexible proof strategy to prove programs correct
% over various isolation variants. For example, programmers can
% \emph{define} isolation guarantees of their choice (by defining $\I$
% appropriately) and then prove programs correct assuming the guarantees
% hold.  The soundness of strengthening $\I$ ensures that a program can
% be safely executed on any system that offers isolation guarantees
% at least as strong as those assumed.

@@

\begin{definition}[\bfseries Interleaved step and multi-step relations]
An interleaved step relation interleaves thread-local reductions with
interference from concurrent threads captured as the rely relation
($R$).  It is defined thus:\vspace*{-10pt}

\begin{smathpar}
\begin{array}{lcl}
\I \vdash (c,\E) \rstepsto (c',\E') & \defeq & \I \vdash 
  (c,\E) \stepsto (c',\E') \\
  &   & \disj (c' = c \conj R(\E, \E') \conj \I(\E'))\\
\end{array}
\end{smathpar}

>>>>

Two rules of the transaction-local RG judgment are shown in
Fig.~\ref{fig:rg-rules}. The rule \rulelabel{RG-Update} is
illustrative of the RG rules for SQL statements; they basically
reflect the structure of the corresponding reduction rule from
Fig.~\ref{fig:txnimp}, and require no further explanation. The rule
\rulelabel{RG-Foreach} defines the RG judgment for a \C{FOREACH} loop.
As is characteristic of loops, the reasoning is pivoted on a loop
invariant $\I$ (not same as the high-level invariant), that needs to
be stable w.r.t $\R$. $I$ must follow $P$, the pre-condition of
\C{FOREACH}, when no elements have been iterated, i.e, when
$y=\emptyset$. The body of the loop can assume the loop invariant, and
the fact that $z$ is an element from the set $x$ being iterated, to
prove its post-condition $Q_c$. Operational semantics ensures that $z$
is added to $y$ at the end of the iteration, hence $Q_c \conj z\in y$
is valid at the end of the iteration, from which $I$ must follow. When
\C{FOREACH} has finished execution, $y$, the set of iterated items, is
the entire set $x$. Thus $I \conj y=x$ must imply the post-condition
$Q$, which also needs to be stable. The rules for conditionals,
sequencing etc., are more-or-less standard, hence elided.

@@

% \begin{definition}[\bfseries Interleaved step and multi-step relations]
% An interleaved step relation interleaves transaction-local reductions with
% legal interference from concurrent transactions as defined by the
% $\I$-constrained rely relation ($\R$).  It is defined thus:
% \begin{smathpar}
% \begin{array}{lcl}
% (c,\stl,\stg) \rstepsto (c',\stl',\stg') & \defeq & \stg \vdash 
%   (c,\stl) \stepsto (c',\stl') \conj \stg'=\stg \\
%   &   & \disj (c' = c \conj \stl'=\stl \conj \R(\stl, \stg, \stg'))\\
% \end{array}
% \end{smathpar}

\begin{definition}[\bfseries Interleaved step and multi-step relations]
An interleaved step relation interleaves a transaction's reduction with
interference from concurrent transactions as captured by the rely
relation. It is defined thus:
\begin{smathpar}
\begin{array}{lcl}
(c,\stg) \rstepsto (c',\stg') & \defeq &  
  (c,\stg) \stepsto (c',\stg') \disj (c' = c \conj R(\stg, \stg'))\\
\end{array}
\end{smathpar}
